```{r setup, include=FALSE}
opts_chunk$set(tidy = FALSE) # preserve wrapped code
opts_chunk$set(dev = 'pdf') # size plots better
require(plyr)
require(ggplot2)
```

# Introduction

## The Problem

### Tail Wagging the Dog

### The Scale of the Problem
* Formal Documents: Journal articles, books, book chapters, theses, consulting reports, etc.
* Informal documents: preliminary analyses, statistical homework
* Online content: web pages, blog posts, forum posts
* Browser metaphor versus page/slide-based metaphor


## The Solution

### Put it in an Email
* Written instructions
* Toolchains
* Open Source
* Workflow

### Toolchain: Data Analysis
1.  Database
1.  Programming language
1.  Extentions to the programming language
    * convenience functions
    * forecasting
    * charts
    * maps
    * interactive charts
    * presentation
    * notebooks

### Toolchain: Presentation

### Open Source
* Documentation
* Support
* Extensibility

### Workflow
LCFD -- Least Commonly Fouled-up Data
* L = Load
* C = Clean
* F = Functions
* D = Do

## Demo -- Database

### Database (SQLite)

## Demo -- Programming

### Basic Programming
```{r}
1 + 1
mean(c(1, 3, 5, 8, 10))
```

### Load/Visualize Data

```{r}
head(baseball[, c("id", "year"  
                  , "stint", "team", "rbi")])
```

### Manipulate Data
```{r}
rbi <- ddply(baseball, .(year), summarise,  
  mean_rbi = mean(rbi, na.rm = TRUE))
head(rbi)
```

### Basic Plots
```{r meanRbiByYear, fig.width = 4, fig.height = 2.25, fig.cap = "mean rbi by year"}
qplot(year, mean_rbi, data = rbi, geom = "line")
```

### Links

* knitr: <http://yihui.name/knitr/>
* R Studio: <http://rstudio.org/>
* R Markdown with R Studio: <http://rstudio.org/docs/authoring/using_markdown>
* My Posts  <http://jeromyanglim.blogspot.com.au/search/label/reproducible%20research>

#### Places to ask questions
* R on StackOverflow: <http://stackoverflow.com/questions/tagged/r>
* LaTeX: <http://tex.stackexchange.com/>
* knitr: <https://github.com/yihui/knitr/issues>

### Thank You

\begin{center}
\LARGE{Questions?}
\end{center}


